---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(dplyr)
library(tidyverse)
library(janitor)
library(tidyr)
library(lme4)
```
```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
system.time(prelim_results <- read.csv("12022100_results.csv", header = TRUE))

prelim_results <- prelim_results %>%
  clean_names()

prelim_results[prelim_results == ""] <- NA

colnames(prelim_results)
head(prelim_results)
```
```{r}
prelim_results <- prelim_results %>%
  mutate(across(c(awake_hour, awake_min, rem_hour, rem_min, light_hour, light_min, deep_hour, deep_min), ~replace_na(., 0))) %>%
  mutate(AwakeNumHour = awake_hour + (awake_min / 60)) %>%
  relocate(AwakeNumHour, .after = awake_min) %>%
  mutate(REMNumHour = rem_hour + (rem_min / 60)) %>%
  relocate(REMNumHour, .after = rem_min) %>%
  mutate(LightNumHour = light_hour + (light_min / 60)) %>%
  relocate(LightNumHour, .after = light_min) %>%
  mutate(DeepNumHour = deep_hour + (deep_min / 60)) %>%
  relocate(DeepNumHour, .after = deep_min) %>%
  mutate(TotalNumHour = AwakeNumHour + REMNumHour + LightNumHour + DeepNumHour) %>%
  relocate(TotalNumHour, .after = DeepNumHour)
```

```{r}
prelim_results <- prelim_results %>%
  group_by(name) %>%
  mutate(participant_id = cur_group_id()) %>%
  ungroup() %>%
  relocate(participant_id, .after = name)

participants_to_exclude <- c("Clarissa Yang", "Zhigui Lee", "Neal Hughes", "Jessica Chen", "DROP")
prelim_results <- prelim_results %>%
  filter(!name %in% participants_to_exclude)
```



```{r}
# fit the linear mixed-effects model to deep sleep hours
model1 <- lmer(DeepNumHour ~ steps + (1 | name), data = prelim_results )
summary(model1)
# fit the linear mixed-effects model to overall sleep hours
model2 <- lmer(TotalNumHour ~ steps + (1 | name), data = prelim_results)
summary(model2)
```

```{r}
# correlation
cor_results <- prelim_results %>%
  select(steps, TotalNumHour, DeepNumHour) %>%
  cor(use = "complete.obs")

print(cor_results)

# multiple linreg
model3 <- lm(DeepNumHour ~ steps, data = prelim_results)
cat("multiple linear regression on number of step count and deep sleep hours- unclustered individual")
summary(model3)

# ANOVA
anova_results <- aov(DeepNumHour ~ steps, data = prelim_results)
cat("ANOVA to look at number of step count deep sleep hours- unclustered individual")
summary(anova_results)

# linear mixed effects, with interaction
model4 <- lmer(DeepNumHour ~ steps * TotalNumHour + (1 | name), data = prelim_results)
summary(model4)
```

Correlation Analysis: The Pearson correlation coefficient between steps and DeepNumHour is 0.0185- a very weak positive linear relationship. Correlation between TotalNumHour and DeepNumHour is 0.2267- likely a weak positive linear relationship.

Multiple Linear Regression: coefficient for steps is 2.804e-06 with a p-value of 0.82 - relationship between steps and DeepNumHour is not statistically significant. R-squared value is 0.0003421- steps explains very little of the variance in DeepNumHour.

ANOVA: F-statistic is 0.052 with a p-value of 0.82, we are unable to reject the null hypothesis on the 0.05 significance level and would like to suggest that there is no significant difference in DeepNumHour across different levels of steps.

Linear mixed-effects with interaction: The model includes interaction terms to see if the relationship between steps and DeepNumHour changes at different levels of TotalNumHour. The coefficients for steps, TotalNumHour, and their interaction are not statistically significant, meaning there is no significant relationship between these variables and DeepNumHour.

------------------------------
#Below is for poster graphing


```{r}
descriptive_stats <- prelim_results %>%
  summarise(
    Mean_Steps = mean(steps, na.rm = TRUE),
    SD_Steps = sd(steps, na.rm = TRUE),
    Min_Steps = min(steps, na.rm = TRUE),
    Max_Steps = max(steps, na.rm = TRUE),
    Mean_DeepNumHour = mean(DeepNumHour, na.rm = TRUE),
    SD_DeepNumHour = sd(DeepNumHour, na.rm = TRUE),
    Min_DeepNumHour = min(DeepNumHour, na.rm = TRUE),
    Max_DeepNumHour = max(DeepNumHour, na.rm = TRUE)
  )

print(descriptive_stats)
```
```{r}
group_stats <- prelim_results %>%
  group_by(study_week) %>%
  summarise(
    Mean_Steps = mean(steps, na.rm = TRUE),
    SD_Steps = sd(steps, na.rm = TRUE),
    Mean_DeepNumHour = mean(DeepNumHour, na.rm = TRUE),
    SD_DeepNumHour = sd(DeepNumHour, na.rm = TRUE),
    Count = n()
  )

print(group_stats)
```
```{r}
# histogram of Steps
ggplot(prelim_results, aes(x = steps)) +
  geom_histogram(binwidth = 1000, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Steps", x = "Number of Steps", y = "Frequency") +
  theme_minimal()
```
```{r}
# density plot- deep sleep hours
ggplot(prelim_results, aes(x = DeepNumHour)) +
  geom_density(fill = "skyblue") +
  labs(title = "Density of Deep Sleep Hours", x = "Deep Sleep Hours", y = "Density") +
  theme_minimal()

ggplot(prelim_results, aes(x = REMNumHour)) +
  geom_density(fill = "skyblue") +
  labs(title = "Density of REM Sleep Hours", x = "REM Sleep Hours", y = "Density") +
  theme_minimal()

ggplot(prelim_results, aes(x = LightNumHour)) +
  geom_density(fill = "skyblue") +
  labs(title = "Density of Core/Light Sleep Hours", x = "Core/Light Sleep Hours", y = "Density") +
  theme_minimal()

ggplot(prelim_results, aes(x = AwakeNumHour)) +
  geom_density(fill = "skyblue") +
  labs(title = "Density of Awake Hours", x = "Awake Hours", y = "Density") +
  theme_minimal()
```
```{r}
prelim_results$Date <- as.Date(prelim_results$activity_sleep_date, format = "%m/%d/%y")

# steps line plot
participants_to_exclude <- c("Clarissa Yang", "Zhigui Lee", "Neal Hughes", "Jessica Chen", "DROP")
prelim_results <- prelim_results %>%
  filter(!name %in% participants_to_exclude)
ggplot(prelim_results, aes(x = Date, y = steps, group = treatment, color = treatment)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Steps Over Time, Faceted by Treatment Week",
       x = "Date",
       y = "Step count") +
      facet_wrap(~ study_week, scales = "free_y") +
       theme(legend.position = "bottom") +
  theme_minimal() +
  theme(legend.position = "bottom")

```

```{r}
prelim_results$Date <- as.Date(prelim_results$activity_sleep_date, format = "%m/%d/%y")

# steps line plot
participants_to_exclude <- c("Clarissa Yang", "Zhigui Lee", "Neal Hughes", "Jessica Chen", "DROP")
prelim_results <- prelim_results %>%
  filter(!name %in% participants_to_exclude)
ggplot(prelim_results, aes(x = Date, y = TotalNumHour, group = treatment, color = treatment)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Sleep Hours Over Time, Faceted by Treatment Week",
       x = "Date",
       y = "Total Hours of Sleep") +
      facet_wrap(~ study_week, scales = "free_y") +
       theme(legend.position = "bottom") +
  theme_minimal() +
  theme(legend.position = "bottom")

```



```{r}
# deep sleep line plot
ggplot(prelim_results, aes(x = Date, y = steps, group = name, color = name)) +
  geom_line() +
  labs(title = "Steps Over Time by Participant",
       x = "Date",
       y = "Step Count") +
  theme_minimal() +
  theme(legend.position = "right")
```
```{r}

# steps faceted by participant
ggplot(prelim_results, aes(x = Date, y = steps)) +
  geom_line(color = "blue") +
  facet_wrap(~ participant_id, scales = "free_y") +
  labs(title = "Steps Over Time by Participant",
       x = "Date",
       y = "Steps (in thousands)") +
  theme_minimal()

# deep sleep hours faceted by participant
ggplot(prelim_results, aes(x = Date, y = DeepNumHour)) +
  geom_line(color = "green") +
  facet_wrap(~ participant_id, scales = "free_y") +
  labs(title = "Deep Sleep Hours Over Time by Participant",
       x = "Date",
       y = "Deep Sleep Hours") +
  theme_minimal()

# scatter plot with deep sleep hours~step regression, faceted by participant
ggplot(prelim_results, aes(x = steps, y = DeepNumHour)) +
  geom_point(color = "blue", alpha = 0.5) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  facet_wrap(~ participant_id, scales = "free_y") +
  labs(title = "Correlation Between Deep Sleep Hours and Steps by Participant",
       x = "Steps",
       y = "Deep Sleep Hours") + theme_minimal()
```
```{r}
# scatter plot with total sleep hours~step regression, faceted by participant
ggplot(prelim_results, aes(x = steps, y = TotalNumHour)) +
  geom_point(color = "blue", alpha = 0.5) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  facet_wrap(~ participant_id, scales = "free_y") +
  labs(title = "Correlation Between Total Sleep Hours and Steps by Participant",
       x = "Steps",
       y = "Total Sleep Hours") + theme_minimal()
```




*** Notes from Dec05 office hour ***
attrition vs. not attrition
1st stage - did the treatment increase steps
regress if you were in treatment
2nd stage - did the steps increase sleep duration
fitted values from 1st stage on the outcome of interest (sleep duration/sleep quality)
AER (applied econometrics with R); function within is ivreg AND panel linear model

Report bootstrap if available
more complex modeling approach: use ‘plm’ https://cran.r-project.org/web/packages/plm/vignettes/A_plmPackage.html
use ‘plm’ function
within this function, there are arguments you can pass to get a iv panel model
run models using outcome for sleep duration and the sleep quality rate
caveat that quality is subjective and noisy because people are bad evaluators of their sleep
more complicated short panel model
prediction of steps based on treatment will be “too precise” because we have no variation within a subject; we need to “blow up” standard errors
the point estimate will be correct but the standard errors will be too small
panel linear model
Approach:
fit instrumental variables model first (ivreg) from AER
standard errors too small
check back-in with Alex once you do these two things
then try doing the panel linear models
option 3: fit regression using ivreg, then use bootstrapping
Alex wants to know if treatment increases step and if treatment increases sleep
2a and 2b from poster
first step: plot data
plot steps first, then duration; plot these side by side
```{r}
library(dplyr)
library(AER)
library(plm)
```

# fit a linear regression model to predict steps based on treatment
treatment_model <- lm(steps ~ treatment, data = prelim_results)
prelim_results$fitted_steps <- fitted(treatment_model)

# regress the fitted steps
treatment_model <- lm(steps ~ treatment, data = prelim_results)
summary(treatment_model)

# instrumental variables regression to regress sleep duration on fitted steps
iv_model <- ivreg(TotalNumHour ~ fitted_steps | treatment, data = prelim_results)
summary(iv_model)


### Linear regression on steps ~ treatment ###

- average number of steps for the control group (treatment = 0) is 6988.
- treatment coefficient: The treatment group (treatment = 1) has an average of 1297.6 more steps than the control group.
- p-value for the treatment coefficient is 0.048, which is less than our common threshold of 0.05, indicating that the treatment effect on steps is statistically significant at the 5% level.
- for R-squared test, our model is only able to explain 2.42% of the variance in steps, which is low....... (but we will have an explanation for this)
- overall linear model on steps vs treatment is significant with a p-value of 0.048.

### Instrumental variable regression with AER on total hours of sleep ~ (linear fitted_steps | treatment) ###

- baseline sleep duration is 4.7 hours.
- **fitted** steps coefficient: for each additional step predicted by the treatment, sleep duration increases by 0.000404 hours.
- p-value for the fitted steps coefficient is 0.023- effect of steps on sleep duration is statistically significant at the usual 0.05 level.
- as for R-squared, our model only explains 3.17% of the variance in total sleep hours

### Overall ###
- treatment significantly increases the number of steps (participants, clustered) by approximately 1297.6 steps.
- increase in steps, as we see being impacted by the treatment, also significantly increases sleep duration, but the effect size is very small (0.000404 hours per step).
Overall, we have statistically significant effects on both models, but the significance (effect size) is really small - our models could explain a very small portion of the variance in the dependent variables.

```{r}
# fit linear regression model to predict steps based on treatment
treatment_model <- lm(steps ~ treatment, data = prelim_results)
prelim_results$fitted_steps <- fitted(treatment_model)

# regress the fitted steps on the outcome of sleep duration
outcome_model <- lm(TotalNumHour ~ fitted_steps, data = prelim_results)
summary(outcome_model)
```

#### Instrumental Variables Regression (`ivreg`) ####

- we use the ivreg function to estimate the coefficients of a linear regression model when there are other dependent variables.
- ivreg also looks at variables that are correlated with the dependent predictors but uncorrelated with errors.
