---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(dplyr)
library(tidyverse)
library(janitor)
library(tidyr)
library(lme4)
```
```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
system.time(prelim_results <- read.csv("12022100_results.csv", header = TRUE))

prelim_results <- prelim_results %>%
  clean_names()

prelim_results[prelim_results == ""] <- NA

colnames(prelim_results)
head(prelim_results)
```
```{r}
prelim_results <- prelim_results %>%
  mutate(across(c(awake_hour, awake_min, rem_hour, rem_min, light_hour, light_min, deep_hour, deep_min), ~replace_na(., 0))) %>%
  mutate(AwakeNumHour = awake_hour + (awake_min / 60)) %>%
  relocate(AwakeNumHour, .after = awake_min) %>%
  mutate(REMNumHour = rem_hour + (rem_min / 60)) %>%
  relocate(REMNumHour, .after = rem_min) %>%
  mutate(LightNumHour = light_hour + (light_min / 60)) %>%
  relocate(LightNumHour, .after = light_min) %>%
  mutate(DeepNumHour = deep_hour + (deep_min / 60)) %>%
  relocate(DeepNumHour, .after = deep_min) %>%
  mutate(TotalNumHour = AwakeNumHour + REMNumHour + LightNumHour + DeepNumHour) %>%
  relocate(TotalNumHour, .after = DeepNumHour)
```

```{r}
prelim_results <- prelim_results %>%
  group_by(name) %>%
  mutate(participant_id = cur_group_id()) %>%
  ungroup() %>%
  relocate(participant_id, .after = name)

participants_to_exclude <- c("Clarissa Yang", "Zhigui Lee", "Neal Hughes", "Jessica Chen", "DROP")
prelim_results <- prelim_results %>%
  filter(!name %in% participants_to_exclude)
```



```{r}
# fit the linear mixed-effects model to deep sleep hours
model1 <- lmer(DeepNumHour ~ steps + (1 | name), data = prelim_results )
summary(model1)
# fit the linear mixed-effects model to overall sleep hours
model2 <- lmer(TotalNumHour ~ steps + (1 | name), data = prelim_results, control = check.scaleX)
summary(model2)
```

```{r}
# correlation
cor_results <- prelim_results %>%
  select(steps, TotalNumHour, DeepNumHour) %>%
  cor(use = "complete.obs")

print(cor_results)

# multiple linreg
model3 <- lm(DeepNumHour ~ steps, data = prelim_results)
cat("multiple linear regression on number of step count and deep sleep hours- unclustered individual")
summary(model3)

# ANOVA
anova_results <- aov(DeepNumHour ~ steps, data = prelim_results)
cat("ANOVA to look at number of step count deep sleep hours- unclustered individual")
summary(anova_results)

# linear mixed effects, with interaction
model4 <- lmer(DeepNumHour ~ steps * TotalNumHour + (1 | name), data = prelim_results)
summary(model4)
```

Correlation Analysis: The Pearson correlation coefficient between steps and DeepNumHour is 0.0185- a very weak positive linear relationship. Correlation between TotalNumHour and DeepNumHour is 0.2267- likely a weak positive linear relationship.

Multiple Linear Regression: coefficient for steps is 2.804e-06 with a p-value of 0.82 - relationship between steps and DeepNumHour is not statistically significant. R-squared value is 0.0003421- steps explains very little of the variance in DeepNumHour.

ANOVA: F-statistic is 0.052 with a p-value of 0.82, we are unable to reject the null hypothesis on the 0.05 significance level and would like to suggest that there is no significant difference in DeepNumHour across different levels of steps.

Linear mixed-effects with interaction: The model includes interaction terms to see if the relationship between steps and DeepNumHour changes at different levels of TotalNumHour. The coefficients for steps, TotalNumHour, and their interaction are not statistically significant, meaning there is no significant relationship between these variables and DeepNumHour.

------------------------------
#Below is for poster graphing


```{r}
descriptive_stats <- prelim_results %>%
  summarise(
    Mean_Steps = mean(steps, na.rm = TRUE),
    SD_Steps = sd(steps, na.rm = TRUE),
    Min_Steps = min(steps, na.rm = TRUE),
    Max_Steps = max(steps, na.rm = TRUE),
    Mean_DeepNumHour = mean(DeepNumHour, na.rm = TRUE),
    SD_DeepNumHour = sd(DeepNumHour, na.rm = TRUE),
    Min_DeepNumHour = min(DeepNumHour, na.rm = TRUE),
    Max_DeepNumHour = max(DeepNumHour, na.rm = TRUE)
  )

print(descriptive_stats)
```
```{r}
group_stats <- prelim_results %>%
  group_by(study_week) %>%
  summarise(
    Mean_Steps = mean(steps, na.rm = TRUE),
    SD_Steps = sd(steps, na.rm = TRUE),
    Mean_DeepNumHour = mean(DeepNumHour, na.rm = TRUE),
    SD_DeepNumHour = sd(DeepNumHour, na.rm = TRUE),
    Count = n()
  )

print(group_stats)
```
```{r}
# histogram of Steps
ggplot(prelim_results, aes(x = steps)) +
  geom_histogram(binwidth = 1000, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Steps", x = "Number of Steps", y = "Frequency") +
  theme_minimal()
```
```{r}
# density plot- deep sleep hours
ggplot(prelim_results, aes(x = DeepNumHour)) +
  geom_density(fill = "skyblue") +
  labs(title = "Density of Deep Sleep Hours", x = "Deep Sleep Hours", y = "Density") +
  theme_minimal()

ggplot(prelim_results, aes(x = REMNumHour)) +
  geom_density(fill = "skyblue") +
  labs(title = "Density of REM Sleep Hours", x = "REM Sleep Hours", y = "Density") +
  theme_minimal()

ggplot(prelim_results, aes(x = LightNumHour)) +
  geom_density(fill = "skyblue") +
  labs(title = "Density of Core/Light Sleep Hours", x = "Core/Light Sleep Hours", y = "Density") +
  theme_minimal()

ggplot(prelim_results, aes(x = AwakeNumHour)) +
  geom_density(fill = "skyblue") +
  labs(title = "Density of Awake Hours", x = "Awake Hours", y = "Density") +
  theme_minimal()
```
```{r}
prelim_results$Date <- as.Date(prelim_results$activity_sleep_date, format = "%m/%d/%y")

# steps line plot
participants_to_exclude <- c("Clarissa Yang", "Zhigui Lee", "Neal Hughes", "Jessica Chen", "DROP")
prelim_results <- prelim_results %>%
  filter(!name %in% participants_to_exclude)
ggplot(prelim_results, aes(x = Date, y = steps, group = name, color = name)) +
  geom_line() +
  labs(title = "Steps Over Time by Participant",
       x = "Date",
       y = "Steps (in thousands)") +
  theme_minimal() +
  theme(legend.position = "right")

```
```{r}
# deep sleep line plot
ggplot(prelim_results, aes(x = Date, y = DeepNumHour, group = name, color = name)) +
  geom_line() +
  labs(title = "Deep Sleep Hours Over Time by Participant",
       x = "Date",
       y = "Deep Sleep Hours") +
  theme_minimal() +
  theme(legend.position = "right")
```
```{r}

# steps faceted by participant
ggplot(prelim_results, aes(x = Date, y = steps)) +
  geom_line(color = "blue") +
  facet_wrap(~ participant_id, scales = "free_y") +
  labs(title = "Steps Over Time by Participant",
       x = "Date",
       y = "Steps (in thousands)") +
  theme_minimal()

# deep sleep hours faceted by participant
ggplot(prelim_results, aes(x = Date, y = DeepNumHour)) +
  geom_line(color = "green") +
  facet_wrap(~ participant_id, scales = "free_y") +
  labs(title = "Deep Sleep Hours Over Time by Participant",
       x = "Date",
       y = "Deep Sleep Hours") +
  theme_minimal()

# scatter plot with deep sleep hours~step regression, faceted by participant
ggplot(prelim_results, aes(x = steps, y = DeepNumHour)) +
  geom_point(color = "blue", alpha = 0.5) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  facet_wrap(~ participant_id, scales = "free_y") +
  labs(title = "Correlation Between Deep Sleep Hours and Steps by Participant",
       x = "Steps",
       y = "Deep Sleep Hours") + theme_minimal()
```
```{r}
# scatter plot with total sleep hours~step regression, faceted by participant
ggplot(prelim_results, aes(x = steps, y = TotalNumHour)) +
  geom_point(color = "blue", alpha = 0.5) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  facet_wrap(~ participant_id, scales = "free_y") +
  labs(title = "Correlation Between Total Sleep Hours and Steps by Participant",
       x = "Steps",
       y = "Total Sleep Hours") + theme_minimal()
```
